Spitzer, Gaia and the Potential of the Milky Way

### 2013-06-19 docs/paper1/paper.tex
first block of text from me

\documentclass{emulateapj}
%\documentclass[preprint]{aastex}
%\documentclass{aastex}
%\usepackage{emulateapj5}
%\usepackage{lscape}
\usepackage{natbib}

@@

\begin{figure}
\plotone{figs/data.eps}
\caption{Summary of data sensitivity: $\Delta d$, $\Delta \mu$ and $\Delta v_{\rm tan}$ as function of $d$.
{\it Also put Sgr and Orphan characteristics on this plot?}
\label{data.fig}
}
\end{figure}

@@

\begin{figure}
\plotone{figs/sgr.eps}
\caption{
\label{sgr.fig}
}
\end{figure}

@@

\begin{figure}
\plotone{figs/boots.eps}
\caption{
\label{boots.fig}
}
\end{figure}

@@

>>>>

%\documentclass{emulateapj}
\documentclass[preprint]{aastex}
%\documentclass{aastex}
%\usepackage{emulateapj5}
%\usepackage{lscape}
\usepackage{natbib}
\usepackage{amssymb,amsmath}

% Convenience
\newcommand{\bs}{\boldsymbol}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}

@@

%% \begin{figure}
%% \plotone{figs/data.eps}
%% \caption{Summary of data sensitivity: $\Delta d$, $\Delta \mu$ and $\Delta v_{\rm tan}$ as function of $d$.
%% {\it Also put Sgr and Orphan characteristics on this plot?}
%% \label{data.fig}
%% }
%% \end{figure}

@@

%% \begin{figure}
%% \plotone{figs/sgr.eps}
%% \caption{
%% \label{sgr.fig}
%% }
%% \end{figure}

@@

%% \begin{figure}
%% \plotone{figs/boots.eps}
%% \caption{
%% \label{boots.fig}
%% }
%% \end{figure}

@@

% Above here is from KVJ. Below is APW.

\section{Method}
With access to full phase-space information for stars in an associated
debris structure, it is no longer necessary to model the full orbit of
the stream. Instead, it is possible to use the individual stars as
potential measures by exploting the simple fact that the stars must
have come from the same progenitor. If the present-day orbital
parameters of the progenitor system are known, this places an
intuitive constraint on the potential of the Milky Way: if the orbits
of the stars and the orbit of the progenitor are integrated
\emph{backwards} in time, the stars should `recombine' with the
progenitor. Conversely, if the potential is incorrect, the orbits of
the stars will diverge from that of the progenitor and thus will not
be recaptured by the satellite system.

The starting point for applying this method is a sample of stars with
known full space kinematics $(\bs{x}_i, \bs{v}_i)|_{t=0}$, the orbital
parameters for the progenitor system $(\bs{x}_p, \bs{v}_p)|_{t=0}$,
and a functional form for the potential, $\Phi$\footnote{Future work
  will investigate a more general ... basis function blah blah.}. For
a given set of potential parameters, the orbits of the stars and
progenitor are integrated backwards for some time (determined from an
estimate of the interaction time of the system). [APW: explain that
  this is not Nbody, just test particles!] At each timestep $t_j$, for
each particle $i$, a set of normalized, relative phase-space
coordinates, $(\bs{q}_i(t_j), \bs{p}_i(t_j))$, are computed using the
instantaneous tidal radius, escape velocity, and phase-space position
of the progenitor. At a particular timestep, these coordinates are
defined as
\begin{equation}
  \bs{q}_i = \frac{\bs{x}_i -
    \bs{x}_p}{R_{tide}}\,\,\,\,\,\,\,\,\bs{p}_i = \frac{\bs{v}_i -
    \bs{v}_p}{v_{esc}}.
\end{equation}
These definitions require an estimate of the mass of the satellite,
$m_{sat}$, which, combined with the computed enclosed mass of the
potential, $m_{enc}$, sets the instantaneous tidal radius and escape
velocity,
\begin{equation}
  R_{tide}=\Big(\frac{m_{sat}}{3m_{enc}}\Big)^{1/3}\,\,\,\,\,\,\,\,
  v_{esc}=\sqrt{\frac{2Gm_{sat}}{R_{tide}}}.
\end{equation}

Qualitatively, when $D_{ps}=\sqrt{|\bs{q}_i|^2+|\bs{p}_i|^2} <
\sqrt{2}$, the star is probably recaptured by the satellite, however
this is not imposed as a hard constraint. Instead, we construct a 6D
distribution of the minimum phase-space vectors for each star over the
individual, integrated orbits. For each star, $i$, the phase-space
distance, $D_{ps}$, is computed at each timestep $t_j$, and the vector
with the minimum phase-space distance is stored
\begin{align}
  t^*_i &= \argmin_{t} D_{ps,i}\\
  A_{ik} &= (\bs{q}_i(t^*_i),\bs{p}_i(t^*_i))
\end{align}
where $k\in\{1,6\}$.

% Discussion?
One of the assumptions of orbit fitting methods for measuring the
potential is that the stars in a given stream are on the same
orbit. \citep{binney??} show that this assumption leads to biased
potential parameters, even without realistic observational errors on
the kinematic information. Conversely, the method presented above is
free of this assumption and instead \emph{relies} on the fact that
each star is on an independent orbit. The power of this method is in
the joint constraint that the best potential is the one that gets the
most of these individual orbits to cross the orbit of the progenitor
over the history of the system.

% APW Note: The optimization, in a way, is getting a bunch of orbits (curves) to intersect another curve (the progenitor orbit) at some points (over time).
